---
title: "flight_delay"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flight_delay}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lab7package)
```


# 1. Loading the flight and weather data

First, the data has been loaded and useful variables have been selected. Furthermore the two datasets have been join together using `origin`, `month`, `day`, `hour` variables in both `weather` and `flight` dataset. All the missing values also have been removed.

```{r}
data("flights", package = "nycflights13")
data("weather", package = "nycflights13")


weather <- weather %>% 
  dplyr::select(!c(wind_gust, year, time_hour))

  #removing some  destinations that have low number of samples
selected_dest <- flights %>% 
  dplyr::group_by(dest) %>%
  dplyr::count() %>% 
  dplyr::filter(n > 50)


dataset <- flights %>% 
  dplyr::filter(dest %in% selected_dest$dest) %>% 
  # joining the weather data
  dplyr::inner_join(weather, c(origin = "origin", month = "month", day = "day", hour = "hour")) %>% 
  # selecting the useful data
  dplyr::select(month, dep_delay, arr_delay, air_time, dest, distance, hour, temp, dewp, humid, wind_dir, wind_speed, precip, pressure, visib) %>%
  # converting month, hour and minute to character so that they will use as factors
  dplyr::mutate(month = as.character(month), hour = as.character(hour)) %>% 
  # scaling the data
  dplyr::mutate(dplyr::across(where(is.numeric) & !dep_delay, scale)) %>% 
  na.omit()


dataset
```



# 2. Splitting Data

Here, the dataset will be split to three different datasets(for training, testing and validating the ridge regression)

```{r}
# 80% of data is going to be used in training dataset
set.seed(1378)
train_index <- caret::createDataPartition(
  dataset$dep_delay,
  p = .8, 
  list = FALSE, 
  times = 1
)

train_data <- dataset[train_index,]
remainder <- dataset[-train_index,]

# Splitting the remainder of the dataset to validation and test sets
set.seed(1378)
test_index <-caret::createDataPartition(
  remainder$dep_delay,
  p = .25, 
  list = FALSE, 
  times = 1
)

test_data <- remainder[test_index, ]
validation_data <- remainder[-test_index, ]

```

# 3. Training Ridge Regression

First we have to find the best value of `lambda` using the `RMSE` statistic on the validation dataset.

```{r}
lambdas <- seq(1, 4, by =0.5)

RMSEs <- purrr::map(lambdas, ~{
  ridge_model <- ridgereg$new(dep_delay~., train_data, lambda = .x)
  
  # calculating RMSE
  y_hat <- ridge_model$predict(validation_data)
  error <- validation_data$dep_delay -  as.vector(y_hat)
  sqrt(mean(error^2))
})
best_lambda <- lambdas[which.min(RMSEs)]
cat("best lambda value is:  ", lambdas[which.min(RMSEs)], "\nRMSE = ", RMSEs[[which.min(RMSEs)]])

```


It seems the value of `lambda` should be `r `best_lambda` to have the best RMSE on the validation set. Now we can a ridge regression with lambda equal to `r best_lambda` and finally predict the  `dep_delay` for the test dataset.

```{r}
ridge_model <- ridgereg$new(dep_delay~., train_data, lambda = best_lambda)
y_hat <- ridge_model$predict(test_data)
```


Now we can calculate the RMSE on the test dataset:

```{r}
error <- test_data$dep_delay -  as.vector(y_hat)
sqrt(mean(error^2))
```

The RMSE on the test data is `r sqrt(mean(error^2))`  which is close to the RMSE on the validation data, and this means the model did not overfit on the train data.
Also we can calculate the $R^2$:

```{r}
error <- train_data$dep_delay -  as.vector(ridge_model$pred())
SSres <- sum(error^2)
SStot <- sum((train_data$dep_delay -  mean(train_data$dep_delay))^2)
R2_train <- 1- SSres/SStot
R2_train

error <- test_data$dep_delay -  as.vector(y_hat)
SSres <- sum(error^2)
SStot <- sum((test_data$dep_delay -  mean(test_data$dep_delay))^2)
R2_test <- 1 - SSres/SStot
R2_test
```


The R squared of the train dataset is `r R2_train` and the R squared of the test dataset is `r R2_test` which means that our model managed to explain the good percentage of the variance of  `dep_delay` variable.
